pirouette: Determining the error BEAST2 makes in inferring a phylogeny

### 2019-03-04 article.tex
Process feedback rsetienne

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{workflow.png}
  \caption{
    \texttt{pirouette} pipeline. 
    The pipeline starts from a true phylogeny (1a), 
    that is converted to an alignment (2a). \rampal{this is unclear; the reader does not know what an experiment is, why there are more, what filtering means, etc}
    The experiments are filtered (3a), possibly based on the
    alignment (2a), to determine which experiments are actually run.
    The experiment that are actually performed, use the 
    alignment (2a) to create a BEAST2 posterior of phylogenies (4a),
    which is compared to the true phylogeny (1a) 
    using the nLTT statistic [\cite{janzen2015approximate}]. (5a) shows the nLTT of the true tree in red,
    the nLTT of all individual posterior trees in grey, and the average
    nLTT statistic of all posterior trees in black.
    Taking the differences between the true tree's nLTT and each posterior's
    nLTT results in an error distribution, here displayed as a histogram
    of error values (6a).
    Optionally, a twin phylogeny (1b) can be generated from a true
    phylogeny (1a), which then results in an error distribution (6b) with
    similar intermediate steps 
  }
  \label{fig:pipeline}
\end{figure}

@@

In the second step selects (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a), the user can specify whether to use the generative model as the inference model or to select the best model from a set of inference models. 
When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Furthermore, because the given phylogeny (on which the alignment is based)
may have followed any tree prior (i.e. speciation model), the user needs
to specify which tree prior is used in the inference. 
When selecting for the best
model, the alignment is used to find the inference model that has the
highest evidence (i.e. marginal likelihood) from a set of candidate inference models.
The evidence of an inference model is estimated using a nested sampling (NS)
approach, as described in \cite{maturana2018model}. The nested sampling is
performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
inference model selection is not available on Windows.\rampal{clarify this. One can use mcbette on a browser, right?}

@@

An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
that will be analyzed in the same way as the true tree.
The twin tree has the same topology as the given phylogeny, 
yet its branching times are simulated according to 
the tree prior used by BEAST2 to produce a posterior 
\giovanni{is it true? is it Yule or BD? Please check}
\richel{It is Yule}.
\rampal{I think it's better to use BD for more generality. Yule is too limited in the trees it can produce.}
The goal of this procedure is to generate a tree 
to use as a control for the main experiment.
We produce the twin tree according to the following procedure.

@@

\verb;pirouette; can be installed easily from CRAN:
\rampal{be careful here; it is not on CRAN yet. Better to say it can be installed from GitHub and will be made available on CRAN)}
\begin{lstlisting}[language=R, floatplacement=H, frame=single]
install.packages("pirouette")
\end{lstlisting}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{workflow.png}
  \caption{
    \texttt{pirouette} pipeline. 
    The pipeline starts from a true phylogeny (1a), 
    that is converted to an alignment (2a).
    \rampal{this is unclear; the reader does not know what an experiment is, why there are more, what filtering means, etc}
    The use supplies one or more experiments, which are a combination of 
    an inference model and the conditions in which it is actually used in
    inference.
    It is determined which experiments will be actually run (3a), 
    possibly based on the
    alignment (2a).
    The experiment that are actually performed, use the 
    alignment (2a) to create a BEAST2 posterior of phylogenies (4a),
    which is compared to the true phylogeny (1a) 
    using the nLTT statistic [\cite{janzen2015approximate}]. (5a) shows the nLTT of the true tree in red,
    the nLTT of all individual posterior trees in grey, and the average
    nLTT statistic of all posterior trees in black.
    Taking the differences between the true tree's nLTT and each posterior's
    nLTT results in an error distribution, here displayed as a histogram
    of error values (6a).
    Optionally, a twin phylogeny (1b) can be generated from a true
    phylogeny (1a), which then results in an error distribution (6b) with
    similar intermediate steps 
  }
  \label{fig:pipeline}
\end{figure}

@@

In the second step selects (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a), the user can specify whether to use the generative model as the inference model or to select the best model from a set of inference models. 
When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Furthermore, because the given phylogeny (on which the alignment is based)
may have followed any tree prior (i.e. speciation model), the user needs
to specify which tree prior is used in the inference. 
When selecting for the best
model, the alignment is used to find the inference model that has the
highest evidence (i.e. marginal likelihood) from a set of candidate inference models.
The evidence of an inference model is estimated using a nested sampling (NS)
approach, as described in \cite{maturana2018model}. The nested sampling is
performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
therefore inference model selection is not available on Windows. Note that
there is the option to do model selection from the browser, 
using \verb;mcbette;. The computation for this model selection is 
performed on the servers of the Travis CI continuous integration service,
that restricts computation time to one hour for free users.

@@

An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
that will be analyzed in the same way as the true tree.
The twin tree has the same topology as the given phylogeny, 
yet its branching times are simulated according to 
the tree prior used by BEAST2 to produce a posterior 
\richel{Should be switched to BD, pirouette Issue 161}.
The goal of this procedure is to generate a tree 
to use as a control for the main experiment.
We produce the twin tree according to the following procedure.

@@

\verb;pirouette; can be installed easily from CRAN:
\rampal{be careful here; it is not on CRAN yet. Better to say it can be installed from GitHub and will be made available on CRAN)}
\richel{We adopted this policy (to say it's on CRAN in advance) with babette as well. Is this a hint to get pirouette on CRAN before the publication of this article :-) ?}
\begin{lstlisting}[language=R, floatplacement=H, frame=single]
install.packages("pirouette")
\end{lstlisting}
